---
title: "01-ACE CTD Profiles"
author: "Lisa-Marie Delpech"
date: "2/23/2022"
output: 
  github_document:
    toc: yes
  html_document:
    toc: yes
    number_sections: yes
    highlight: pygments
    code_folding: 'show'
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Libraries, message = FALSE}
require(tidyverse)
require(viridis)
require(gsw)
require(oce)
require(ocedata)
require(ggoce)
require(cowplot)
require(ggrepel)
```

```{r Load data}
ctd.profile.data <- readRDS("./R_Data/CTD_profile_data.rds") # CTD data concatenated in long format df
head(ctd.profile.data)
length(table(ctd.profile.data$EVENTNBR))
ctd.profile.data$EVENTNBR <- as.factor(ctd.profile.data$EVENTNBR)

meta <- read.table(file = "./Metadata/meta_CTD_ess_V4_4.csv", header=TRUE, sep = ";", dec = ".", stringsAsFactors = TRUE) # Metadata to get event numbers
length(table(meta$Event_number)) # 49 event numbers vs. 61 available CTD profiles

meta.sum <- meta %>% 
  select("Event_number","Depth_m","Latitude","Longitude","Temperature_ITS90","Salinity_PSS78","Density_kg.m2","dO2_µmol.kg", "Water_mass_detailed") %>% 
  group_by(Event_number, Depth_m, Water_mass_detailed) %>% # Remove filter information 
  summarize_if(.predicate = is.numeric, .funs = mean)
```

The README file accompanying the metadata in Henry *et al.* (2020) describes quality flags as follow:

> Quality Flags
Each variable was inspected and quality flags (WHP-Exchange Format, Barna, Swift and Diggs, 2016; https://exchange-format.readthedocs.io/en/latest/) applied using the following conditions:
* Flag 1 Not Calibrated - Oxygen variables (CTD) flagged until comparison with bottle samples or suitable climatology dataset e.g. GLODAP, WOCE or WOA  
* Flag 2 Acceptable Measurement  
* Flag 3 Questionable Measurement - CTDTEMP or CTDSAL failed spike or gradient tests (SeaDataNet Data Quality Control Procedures Version 2.0). BACKSC where > 50 % of values flagged as Flag 4, see below.  
* Flag 4 Bad Measurement - Bad flag applied during Sea-Bird Data Processing Modules, scan variable value is negative number, variable value identified as being anomalous at beginning or end of upcast or downcast during visual inspection (see ace_ctd_visual_inspection.csv for full list of data points identified), values fall outside acceptable sensor specific ranges (see information on sensor specific flagging below).  
* Flag 5 Not Reported - No instances where this was applied. 
* Flag 6 Interpolated over pressure interval larger than 2dbar. 
* Flag 7 Despiked - No instances where this was applied.  
* Flag 9 Not Sampled - Sensor was not connected. 
Flags can be identified from the following header format [PARAMETER]_FLAG_W (WHP Exchange Format).  

We here remove flags 3 and 4 for salinity, temperature and oxygen. 

```{r Filter quality flags}
CTDTMP_QUAL <- ctd.profile.data$CTDTMP
CTDTMP_QUAL[which(ctd.profile.data$CTDTMP_FLAG_W %in% c(3,4))] <- NA
CTDSAL_QUAL <- ctd.profile.data$CTDSAL
CTDSAL_QUAL[which(ctd.profile.data$CTDSAL_FLAG_W %in% c(3,4))] <- NA
CTDOXY_QUAL <- ctd.profile.data$CTDOXY
CTDOXY_QUAL[which(ctd.profile.data$CTDOXY_FLAG_W %in% c(3,4))] <- NA

ctd.profile.filtered <- ctd.profile.data %>% 
  mutate(CTDTMP = CTDTMP_QUAL) %>% # , .after = CTDTMP
  mutate(CTDSAL = CTDSAL_QUAL) %>% # , .after = CTDSAL
  mutate(CTDOXY = CTDOXY_QUAL) # , .after = CTDOXY
```

Check availability of CTD profiles for metadata and metagenomic samples. These are the events not available in in CTD profiles.

```{r Check availability of CTD profiles}
meta.sum[which(!meta.sum$Event_number %in% ctd.profile.filtered$EVENTNBR),]
```

Event 75 is not available as a CTD profile while it is in the metadata. Hence we will not be able to classify this sample's water mass based on the CTD data.
Note that event 317 is not in the metagenomes dataset but included here with all available CTD profiles for the metadata.

# Compute absolute salinity and potential temperature

```{r GSW equations}
ctd <- ctd.profile.filtered %>% 
  mutate(Absolute_Salinity = gsw_SA_from_SP(CTDSAL, CTDPRS, LONGITUDE, LATITUDE)) %>% 
  mutate(Potential_temperature = gsw_pt_from_t(Absolute_Salinity, CTDTMP, CTDPRS, p_ref = 0),
         Conservative_temperature = gsw_CT_from_t(Absolute_Salinity, CTDTMP, CTDPRS))
```

# T-S plots per event

```{r Plot for all events}
# TS plot for all
(plot.TS.all <- ggplot(ctd %>% filter(CASTTYPE == "downcast"),
                       aes(x = CTDSAL, y = Conservative_temperature)) +
   geom_isopycnal(salinity_type = "practical",
                  temperature_type = "conservative",
                  ref_longitude = 0,
                  ref_latitude = 0,
                  breaks = seq(25,28,0.2),
                  linetype = "dotted") +
   geom_path(aes(group = EVENTNBR),
             size = .3,
             col = "#777777") +
   labs(x = "Practical Salinity (psu)", y = "Conservative Temperature (°C)") +
   scale_x_continuous(breaks = seq(32,35,0.2)) +
   scale_y_continuous(breaks = seq(-1,12,2)) +
   theme_bw())

# Depth-T for all
(plot.Depth.T.all <- ggplot(data = ctd %>% filter(CASTTYPE == "downcast",
                                                  DEPTH < 2000),              # Only affects event 1465 for which depth goes over 3000m
                            aes(x = Conservative_temperature, y = DEPTH)) +
    geom_path(aes(group = EVENTNBR),
              size = .3, col = "#777777") +
    scale_x_continuous(breaks = seq(-1,12,2)) +
    scale_y_reverse() +
    labs(x = "Conservative Temperature (°C)", y = "Depth (m)") +
    theme_bw())

# Depth-S for all
(plot.Depth.S.all <- ggplot(data = ctd %>% filter(CASTTYPE == "downcast",
                                                  DEPTH < 2000),
                            aes(x = CTDSAL, y = DEPTH)) +
    geom_path(aes(group = EVENTNBR),
              size = .3, col = "#777777") +
    scale_x_continuous(breaks = seq(32,35,0.4)) +
    scale_y_reverse() +
    labs(x = "Practical Salinity (psu)", y = "Depth (m)") +
    theme_bw())
```

```{r ggsave TS plot, eval = FALSE, include = FALSE}
saveRDS(plot.TS.all, "./R_Data/TS_plot_all.rds")
```


```{r Add highlight for each event, fig.dim = c(9.4, 12.6), warning=FALSE}
for (event in unique(ctd$EVENTNBR)){
  if (event %in% meta.sum$Event_number){
    df.ctd = ctd %>% filter(CASTTYPE == "downcast",
                            ##CTDSAL >= 32 & CTDSAL <= 35,
                            EVENTNBR == event)
    df.meta = meta.sum %>% filter(Event_number == event)
    
    # Add TS plot for one event
    plot.TS <- plot.TS.all +
      geom_path(data = df.ctd,
                aes(x = CTDSAL, y = Conservative_temperature, col = DEPTH),
                size = .7) +
      scale_color_gradient(high = "red", low = "blue") +
      geom_point(data = df.meta,
                 aes(x = Salinity_PSS78, y = Temperature_ITS90, col = Depth_m)) +
      geom_point(data = df.meta,
                 aes(x = Salinity_PSS78, y = Temperature_ITS90),
                 col = "black", fill = "transparent", shape = 21) +
      geom_label_repel(data = df.meta,
                       aes(x = Salinity_PSS78, y = Temperature_ITS90, label = Water_mass_detailed),
                       size = 3,
                       label.padding = 0.15, force_pull = .5,
                       label.r = 0.07) +
      scale_fill_gradient(high = "red", low = "blue", guide = "none") +
      labs(col = "Depth (m)", title = paste("Event number", event)) +
      theme(plot.title = element_text(hjust = 0.5))
    
    # Add Depth-T plot for one event
    plot.Depth.T <- plot.Depth.T.all +
      geom_path(data = df.ctd %>% filter(DEPTH < 2000),
                aes(x = Conservative_temperature, y = DEPTH, col = DEPTH),
                size = .7) +
      geom_point(data = df.meta,
                 aes(x = Temperature_ITS90, y = Depth_m, col = Depth_m)) +
      scale_color_gradient(high = "red", low = "blue") +
      geom_point(data = df.meta,
                 aes(x = Temperature_ITS90, y = Depth_m),
                 col = "black", fill = "transparent", shape = 21) +
      #scale_fill_gradient(high = "red", low = "blue", guide = "none") +
      geom_label_repel(data = df.meta,
                       aes(x = Temperature_ITS90, y = Depth_m, label = paste(Depth_m,"m")),
                       size = 2) +
      labs(col = "Depth (m)")
    
    # Add Depth-S plot for one event
    plot.Depth.S <- plot.Depth.S.all +
      geom_path(data = df.ctd %>% filter(DEPTH < 2000),
                aes(x = CTDSAL, y = DEPTH, col = DEPTH),
                size = .7) +
      scale_color_gradient(high = "red", low = "blue") +
      geom_point(data = df.meta,
                 aes(x = Salinity_PSS78, y = Depth_m, col = Depth_m)) +
      geom_label_repel(data = df.meta,
                       aes(x = Salinity_PSS78, y = Depth_m, label = paste(Depth_m,"m")),
                       size = 2) +
      geom_point(data = df.meta,
                 aes(x = Salinity_PSS78, y = Depth_m), col = "black", fill = "transparent", shape = 21) +
      #scale_fill_gradient(high = "red", low = "blue", guide = "none") +
      labs(col = "Depth (m)") +
      theme(legend.position = "none")
    
    # All together
    plot.Depth.S.T <- plot_grid(plot.Depth.S, plot.Depth.T, rel_widths = c(1,1.3))
    print(
      plot_grid(plot.TS, plot.Depth.S.T, nrow = 2, ncol = 1)
    )
    
    #ggsave(paste0("./R_Plots/CTD_Profiles/CTD_profile_",event,".pdf"), height = 10, width = 8, paper = "a4")
  }
}
```

Note that event 1029 has a limited downcast but the upcast might be useful

```{r Zoom in on TS diagram, fig.dim = c(8,5), warning=FALSE}
# Refine salinity between 33.6 and 34.8
# And temperature between -2 and 5

plot.TS.all.zoom <- ggplot(ctd %>% filter(CASTTYPE == "downcast",
                                          CTDSAL >= 33.6, CTDSAL <= 34.8,
                                          Conservative_temperature >=-2, Conservative_temperature <= 5),
                           aes(x = CTDSAL, y = Conservative_temperature)) +
  geom_isopycnal(salinity_type = "practical",
                 temperature_type = "conservative",
                 ref_longitude = 0,
                 ref_latitude = 0,
                 breaks = seq(26.7,28,0.1),
                 linetype = "dotted") +
  geom_path(aes(group = EVENTNBR),
            size = .3,
            col = "#777777") +
  labs(x = "Practical Salinity (psu)", y = "Conservative Temperature (°C)") +
  scale_x_continuous(breaks = seq(33.6,34.8,0.1)) +
  scale_y_continuous(breaks = seq(-2,5,1)) +
  theme_bw()

for (event in unique(ctd %>% filter(CASTTYPE == "downcast", CTDSAL >= 33.6, CTDSAL <= 34.8, Conservative_temperature >=-2, Conservative_temperature <= 5) %>% pull(EVENTNBR))){
  if (event %in% meta.sum$Event_number){
    
    df.ctd = ctd %>% filter(CASTTYPE == "downcast",
                            CTDSAL >= 33.6, CTDSAL <= 34.8,
                            Conservative_temperature >=-2, Conservative_temperature <= 5,
                            EVENTNBR == event)
    df.meta = meta.sum %>% filter(Event_number == event,
                                  Salinity_PSS78 >= 33.6, Salinity_PSS78 <= 34.8,
                                  Temperature_ITS90 >= -2, Temperature_ITS90 <= 5)
    
    # Add TS plot for one event
    plot.TS.path <- plot.TS.all.zoom +
      geom_path(data = df.ctd,
                aes(x = CTDSAL, y = Conservative_temperature, col = DEPTH),
                size = .7) +
      scale_color_gradient(high = "red", low = "blue") +
      labs(col = "Depth (m)", title = paste("Zoom - Event number", event)) +
      theme(plot.title = element_text(hjust = 0.5))
    
    if (nrow(df.meta) != 0){
      print(
      plot.TS.path +
        geom_point(data = df.meta,
                   aes(x = Salinity_PSS78, y = Temperature_ITS90, col = Depth_m)) +
        geom_point(data = df.meta,
                   aes(x = Salinity_PSS78, y = Temperature_ITS90),
                   col = "black", fill = "transparent", shape = 21) +
        scale_fill_gradient(high = "red", low = "blue", guide = "none") +
        geom_label_repel(data = df.meta,
                         aes(x = Salinity_PSS78, y = Temperature_ITS90, label = paste(Depth_m,"m")),
                         size = 2) +
        labs(col = "Depth (m)", title = paste("Zoom - Event number", event)) +
        theme(plot.title = element_text(hjust = 0.5))
      )
    }
  }
  
  #ggsave(paste0("./R_Plots/CTD_Profiles/TS_zoom_",event,".pdf"), height = 5, width = 8, paper = "a4")
}
```


# Oxygen-Salinity and Oxygen-Temperature plots

In order to determine water masses and remove some ambiguity around some samples, we need to also add salinity oxygen plots to our data.

Notes from the data collection:

> A note on oxygen: During the data conversion process, the ‘tau’ and ‘hysteresis’ corrections were applied to the oxygen data. The ‘tau’ correction improves the response of the measured signal in regions of large oxygen gradients, and also may amplify residual noise in the signal (CalCOFI CTD Data Algorithms). The ‘tau’ correction computes the derivative of the oxygen signal with respect to time with a user-input window size), using a linear regression to determine the slope. The correction was applied using the Data Conversion step where the module uses a window looking backwards in time and a window size of two seconds was defined. Although it is recommended that the ‘tau’ correction be applied during the ‘Derive’ module for more accurate results, a subset of files were processed using that processing order and there was no difference in the derived oxygen values. The hysteresis correction corrects the oxygen voltage values for changes in membrane permeability as pressure varies. Correction coefficients stored in the XMLCON file are applied with calculations based on the current pressure and how long the session spent at previous pressures. Two oxygen variables are provided: (1) dissolved oxygen concentration in micromols/kg and (2) dissolved oxygen saturation, the theoretical saturation limit of the water at the local temperature and salinity value, with local pressure reset to zero (1 atmosphere) (See SBE Data Processing Software Manual for SBE Data Processing Software v7.26.8 & CalCOFI CTD Data Algorithms). Dissolved oxygen saturation is calculated using the equations of Garcia and Gordon (1992) which is based on the equation of Weiss (1970) but reduced the error in cold waters and presents improved performance in waters between -5 to 50 degrees Celsius. 

Regarding the quality of those values, Henry *et al.* report:

> Variables CTDOXY and CTDOXYSAT were flagged with Flag 1 as per the CTD cast data (See section Quality Flags).  
> The oxygen variables have not been corrected further, but should be compared to oxygen bottle data when made available. An initial comparison to World Ocean Atlas 2018 data (WOA18; Figure 3; Garcia *et al.*, 2018) suggests a substantial offset which should be corrected for when using the oxygen data.


```{r Plots dO-S, fig.dim = c(9.4, 12.6), warning = FALSE}
# Oxygen-Salinity diagram for all samples
(plot.dOS.all <- ggplot(ctd %>% filter(CASTTYPE == "downcast"),
                      aes(x = CTDSAL, y = CTDOXY)) +
  geom_path(aes(group = EVENTNBR),
            size = .3,
            col = "#777777") +
  labs(x = "Practical Salinity (psu)", y = "Dissolved Oxygen (µmol/kg)") +
  scale_x_continuous(breaks = seq(32,35,0.2)) +
  scale_y_continuous(breaks = seq(150,400,20)) +
  theme_bw())

#Temperature-Oxygen diagram for all samples
(plot.TdO.all <- ggplot(ctd %>% filter(CASTTYPE == "downcast"),
                        aes(x = CTDOXY, y = Conservative_temperature)) +
  geom_path(aes(group = EVENTNBR),
            size = .3,
            col = "#777777") +
  labs(x = "Dissolved Oxygen (µmol/kg)", y = "Conservative Temperature (°C)") +
  scale_x_continuous(breaks = seq(150,400,20)) +
  scale_y_continuous(breaks = seq(-1,12,2)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)))

#Depth-Oxygen for all samples
(plot.Depth.O2.all <- ggplot(data = ctd %>% filter(CASTTYPE == "downcast",
                                                  DEPTH < 2000),
                             aes(x = CTDOXY, y = DEPTH)) +
  geom_path(aes(group = EVENTNBR),
            size = .3, col = "#777777") +
  scale_x_continuous(breaks = seq(150,400,20)) +
  scale_y_reverse() +
  labs(x = "Dissolved Oxygen (µmol/kg)", y = "Depth (m)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)))


for (event in unique(ctd$EVENTNBR)){
  if (event %in% meta.sum$Event_number){
  #print(event)
  df.ctd = ctd %>% filter(CASTTYPE == "downcast",
                          EVENTNBR == event)
  df.meta = meta.sum %>% filter(Event_number == event)
  
  # Add TS plot for one event
  plot.dOS <- plot.dOS.all +
    geom_path(data = df.ctd,
              aes(x = CTDSAL, y = CTDOXY, col = DEPTH),
              size = .7) +
    scale_color_gradient(high = "red", low = "blue") +
    geom_point(data = df.meta,
               aes(x = Salinity_PSS78, y = dO2_µmol.kg, col = Depth_m)) +
    geom_point(data = df.meta,
               aes(x = Salinity_PSS78, y = dO2_µmol.kg),
               col = "black", fill = "transparent", shape = 21) +
    scale_fill_gradient(high = "red", low = "blue", guide = "none") +
    labs(col = "Depth (m)", title = paste("Event number", event)) +
    theme(plot.title = element_text(hjust = 0.5)) +
    geom_label_repel(data = df.meta,
                     aes(x = Salinity_PSS78, y = dO2_µmol.kg, label = paste(Depth_m,"m")),
                     size = 2)
  
  plot.TdO <- plot.TdO.all +
    geom_path(data = df.ctd,
              aes(x = CTDOXY, y = CTDTMP, col = DEPTH),
              size = .7) +
    scale_color_gradient(high = "red", low = "blue") +
    geom_point(data = df.meta,
               aes(x = dO2_µmol.kg, y = Temperature_ITS90, col = Depth_m)) +
    geom_point(data = df.meta,
               aes(x = dO2_µmol.kg, y = Temperature_ITS90),
               col = "black", fill = "transparent", shape = 21) +
    scale_fill_gradient(high = "red", low = "blue", guide = "none") +
    labs(col = "Depth (m)") +
    theme(plot.title = element_text(hjust = 0.5)) +
    geom_label_repel(data = df.meta,
                     aes(x = dO2_µmol.kg, y = Temperature_ITS90, label = paste(Depth_m,"m")),
                     size = 2) +
    theme(legend.position = "none")
  
  plot.Depth.O2 <- plot.Depth.O2.all +
    geom_path(data = df.ctd %>% filter(DEPTH < 2000),
              aes(x = CTDOXY, y = DEPTH, col = DEPTH),
              size = .7) +
    scale_color_gradient(high = "red", low = "blue") +
    geom_point(data = df.meta,
               aes(x = dO2_µmol.kg, y = Depth_m, col = Depth_m)) +
    geom_label_repel(data = df.meta,
                     aes(x = dO2_µmol.kg, y = Depth_m, label = paste(Depth_m,"m")),
                     size = 2) +
    geom_point(data = df.meta,
               aes(x = dO2_µmol.kg, y = Depth_m), col = "black", fill = "transparent", shape = 21) +
    labs(col = "Depth (m)")
  
  
  # All together
  plot2 <- plot_grid(plot.TdO, plot.Depth.O2, rel_widths = c(1.1,1))
  print(
    plot_grid(plot.dOS, plot2, nrow = 2, ncol = 1)
  )
  #ggsave(paste0("./R_Plots/CTD_Profiles/CTD_Oxygen_",event,".pdf"), height = 11, width = 8, paper = "a4")
  }
}
```

